---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
author: 'Group: M358k-ers'
date: '2020-09-18'
  
---

<!-- The author of this template is Dr. Gordan Zitkovic.-->
<!-- The code chunk below contains some settings that will  -->
<!-- make your R code look better in the output pdf file.  -->
<!-- If you are curious about what each option below does, -->
<!-- go to https://yihui.org/knitr/options/ -->
<!-- If not, feel free to disregard everything ...  -->
```{r echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.align="center",
  fig.pos="t",
  strip.white = TRUE
)
```
<!-- ... down to here. -->

## Problem 1

#### **Problem 1.1.** What is the probability that the gambler wins the game?

The probability of winning is equal to the $P(Win|Side)$, because the gambler has to choose a side.
Probability of winning for any chosen side is equal because we assume we have a fair die.
Thus, we equate $P(Win|Side)$ as the probability that the gambler wins the game

Now, let us define the variables

$Side$: The side that the gambler chooses

$D_1, D_2, D_3$: Each $D_i$ represents the side of Die $i$ landed on

$P(Win| Side) = P(D_1=Side \cap D_2=Side \cap D_3 = Side) + P(D_1=Side \cap D_2=Side \cap D_3 \neq Side)$
$+ P(D_1=Side \cap D_2\neq Side \cap D_3 =Side)  P(D_1\neq Side \cap D_2=Side \cap D_3 = Side)$

Note: The outcomes of the die are independent, so we can change the intersections to multiplication


$P(Win| Side) = P(D_1=Side)P( D_2=Side) P( D_3 = Side) + P(D_1=Side)P(D_2=Side) P( D_3 \neq Side)$
$+ P(D_1=Side)P(D_2\neq Side)P( D_3 =Side) +  P(D_1\neq Side)P(D_2=Side )P( D_3 = Side)$

Note: $P(D_i=Side) = 1/6$ for any $i$ and $P(D_i\neq Side)=5/6$ 

$P(Win | Side) = (1/6)^3+(1/6)^2(5/6)+(1/6)^2(5/6)+(1/6)^2(5/6)$

$\textbf{P(Win | Side) = (1/6)^3 + 3(1/6)^2(5/6) = 0.07407407407}$

```{r}
792/11
```

### Part a.
You can write an explanation here, too. 

```{r}
x=c(2,4)
print(x^2)
```

### Part b.
```{r}

```

### Part e.
```{r}

```
  
### Part f.
```{r}

```
  
  
### Part h.
```{r}

```


## Problem 5.



