---
title: "Project 1"
author: 'Sunny Kharel, William Easterby, Gabby Toutin'
date: '2020-09-18'
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

<!-- The author of this template is Dr. Gordan Zitkovic.-->
<!-- The code chunk below contains some settings that will  -->
<!-- make your R code look better in the output pdf file.  -->
<!-- If you are curious about what each option below does, -->
<!-- go to https://yihui.org/knitr/options/ -->
<!-- If not, feel free to disregard everything ...  -->
```{r echo=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.align="center",
  fig.pos="t",
  strip.white = TRUE
)
```
<!-- ... down to here. -->

## Problem 1

#### **Problem 1.1.** What is the probability that the gambler wins the game?

The probability of winning is equal to the $P(Win|Side)$, because the gambler has to choose a side.
Probability of winning for any chosen side is equal because we assume we have a fair die.
Thus, we equate $P(Win|Side)$ as the probability that the gambler wins the game

Now, let us define the variables

$Side$: The side that the gambler chooses

$D_1, D_2, D_3$: Each $D_i$ represents the side of Die $i$ landed on

$$P(Win| Side) = P(D_1=Side \cap D_2=Side \cap D_3 = Side) + P(D_1=Side \cap D_2=Side \cap D_3 \neq Side)$$
$$+ P(D_1=Side \cap D_2\neq Side \cap D_3 =Side)  P(D_1\neq Side \cap D_2=Side \cap D_3 = Side)$$

Note: The outcomes of the die are independent, so we can change the intersections to multiplication


$$P(Win| Side) = P(D_1=Side)P( D_2=Side) P( D_3 = Side) + P(D_1=Side)P(D_2=Side) P( D_3 \neq Side)$$
$$+ P(D_1=Side)P(D_2\neq Side)P( D_3 =Side) +  P(D_1\neq Side)P(D_2=Side )P( D_3 = Side)$$

Note: $P(D_i=Side) = 1/6$ for any $i$ and $P(D_i\neq Side)=5/6$ 

$$P(Win | Side) = (1/6)^3+(1/6)^2(5/6)+(1/6)^2(5/6)+(1/6)^2(5/6)$$

$$\textbf{P(Win)=P(Win | Side) = (1/6)^3 + 3(1/6)^2(5/6) = 0.07407407407}$$


#### **Problem 1.2** Who has the long-run advantage in the game: the house or the gambler? By how much on the dollar?
Let us examine the probability of earning x dollars given we put in $1

$$P(X = -1) =   P(D_1 \neq Side)P(D_2 \neq Side)P(D_3 \neq Side) = (5/6)^3$$

$$P(X = 1) =   P(D_1 = Side)P(D_2 \neq Side)P(D_3 \neq Side)+ P(D_1 \neq Side)P(D_2 = Side)P(D_3 \neq Side)$$
$$+P(D_1 \neq Side)P(D_2 \neq Side)P(D_3 = Side) = 3(1/6)(5/6)^2$$

$$P(X=2) =   P(D_1 = Side)P(D_2 = Side)P(D_3 \neq Side) + P(D_1 = Side)P(D_2 \neq Side)P(D_3 = Side)$$
$$+ P(D_1 \neq Side)P(D_2 = Side)P(D_3 = Side) = 3(1/6)^2(5/6)$$

$$P(X=3) =   P(D_1 = Side)P(D_2 = Side)P(D_3 = Side) = (1/6)^3$$

The expected value of dollars is:

$$E[X]=-1\$ \cdot (5/6)^3 + 1\$ \cdot 3(1/6)(5/6)^2 + 2\$\cdot 3(1/6)^2(5/6) + 3\$\cdot (1/6)^3 = -0.0787037037$$

To see who has the long run advantage, we take the expected earnings of n successive games, in which we put in 
a dollar. The expected earning of n successive games, E[earnings in n games], equals n*E[X]. E[X] is negative 
and n is positive, so we are always expected to be at a loss with more successive games, with **-8 cents/dollar**.

#### **Problem 1.3** How would you change the amount of the winnings in every round of Chuck-a-Luck to make the game fair? Say that doubles pay $a per $1 bet rather than $2 and triples pay $b per $1 bet in stead of $3. What is the condition on a and b?


For the game to be fair, E[X], as modeled in 1.2, should be equal to 0. Thus with some algebra:

$$E[X]=-1\$ \cdot (5/6)^3 + 1\$ \cdot 3(1/6)(5/6)^2 + a\cdot2\$\cdot 3(1/6)^2(5/6) + b\cdot3\$\cdot (1/6)^3 = 0$$

This reduces to 

$$30\cdot a + b = 50$$

#### **Problem 1.4** Get three regular dice or use software to create simulated values of the outcomes of rounds of Chuck-a-Luck with the original reward system. Play 100 independent rounds of the game, starting with a new dollar each time. What is the average of the gambler’s winnings in these 100 rounds?


```{r}
rounds = rbinom(100, 3, 1/6)
print(paste0("The average is $", mean(rounds)), quote = FALSE)
```

#### **Problem 1.5** Take another look at your answers to problems 1.2 and 1.4. How do they relate to the Strong Law of Large Numbers?

The answer to 1.4 shows that over time the expected winnings (or losings) trend towards the overall expected value in 1.2. Even in just 100 games, there is already a lose of `r (1-mean(rounds))*100` cents on the dollar, and over even more rounds we would expect it to converge to around 8 cents lose per dollar.


#### **Problem 1.6**  Now we are going to look at a particular gambler who arrives at the casino with $2 in his pocket intending to play Chuck-a-Luck. At the beginning of every round of the game, he invests exactly $1 (if he has the money). He continues to play round after round of the game until he either loses everything, or his wealth reaches at least $4.


```{r}
dollars=2
round_count=0
wealth=c(2)
while ((dollars > 0) & (dollars < 4)){
  dollars=dollars-1
  round_count=round_count+1
  dollars=dollars+rbinom(1,3,1/6)
  wealth=append(wealth, dollars)
}
plot(wealth, type = "s")

```




